Computing properties of independent sets by generic programming tensor networks \thanks{\funding{...}}

### 2022-01-05 paper/paper.tex

\section{Introduction}
In graph theory and combinatorial optimization, there are many interesting and hard computational problems concerning various properties of independent sets.
%An independent set of a graph is a subset of its vertices that no two vertices in this set are adjacent to each other.
For an undirected graph $G = (V,E)$, an independent set $I \subseteq V$ is a set of vertices that for any vertex pair $u,v \in I$, $(u, v) \not\in E$. %there is no edge connecting $u$ and $v$ in $G$. 
One of the independent set properties that many computational complexity theorists are most interested in is to find a maximum independent set (MIS) and the size of the MIS, $\alpha(G) \equiv \max_{I}|I|$. 
% The primary interest of many computational complexity theorists is finding the maximum size of such vertex sets $\alpha(G) \equiv \max_{I}|I|$ and one of the sets of this size.
Finding $\alpha(G)$ exactly is a hard computational problem, and it is NP-hard even to approximate $\alpha(G)$ within $|V|^{1-\epsilon}$ ~\cite{Hastad1996},
meaning there is unlikely a polynomial time algorithm to find an independent set with size $\alpha(G)/|V|^{1-\epsilon}$ for an arbitrarily small positive $\epsilon$.
Naive exhaustive search for an MIS requires a computational time of $O\left(2^{|V|} \right)$.
More efficient exact algorithms have been developed for finding MISs, such as the branching algorithms~\cite{Tarjan1977, Robson1986} and dynamic programming based algorithms~\cite{Courcelle1990, Fomin2013}.
The branching algorithms can reduce the base of the exponential time scaling to, e.g., $1.1996^{|V|} {|V|}^{O(1)}$~\cite{Xiao2017}, 
while dynamic programming approaches~\cite{Courcelle1990, Fomin2013} works better for graphs with a small treewidth $\text{tw}(G)$, producing algorithms of complexity $O(2^{\text{tw}(G)} \text{tw}(G) |V|)$.
% People are eager to find better algorithms for solving the MIS problem;
There is immense interest in finding better algorithms to solve the MIS problem, not only because it has 
a wide range of applications in scheduling, logistics, wireless networks and telecommunication, and computer vision, etc.~\cite{Butenko2003, Wu2015}, 
% scheduling, logistics, social network analysis, bioinformatics, wireless networks and telecommunication,
% map labelling, computer vision, etc.~\cite{Butenko2003, Wu2015}, 
but also because it is a well-known NP-complete problem that can be mapped from many other important combinatorial optimization problems in polynomial time such as the 3-satisfiability problem, the maximum clique problem, and the minimum vertex cover problem~\cite{Moore2011}.

%more concretely, an maximal independent set of a graph $G$ corresponds to a maximum clique in the complement graph of $G$,
%and the complement vertex set of an MIS corresponds to a minimum vertex cover of the graph $G$.
%In this paper, the term ``maximal'' has a different meaning from ``maximum''  which we discussed above; a maximal independent set is an independent set that is not a subset of any other independent set, but its size may not be the maximum.
In this paper, we do not limit our study to just finding an MIS and its size. 
We introduce a tensor-network based framework to compute the various \textit{properties} pertaining to independent sets. These properties include, for example, the number and enumeration of independent sets or maximal independent sets of a given size. We map these independent set problems into generic tensor network contraction with specially designed tensor element algebra.
% There are many other interesting problems pertaining to independent sets, such as the number of independent sets of a given size
% and enumerating independent sets at a give size or what we call computing \textit{properties}.
Some of these problems are of great interest in physics applications such as the hard-core lattice gas model~\cite{Dyre2016, Fernandes2007}
in statistical mechanics and the Rydberg hamiltonian with neutral atoms~\cite{Pichler2018, Ebadi2022} \red{replace with experiment paper when ready};
they can, for example, be used to understand phase transitions, to identify harder graphs in an ensemble of graphs~\cite{Ebadi2022} and to analyze the overlap gap property~\cite{Gamarnik2013, Gamarnik2019}.

@@

\section{Tensor networks}
Tensor network~\cite{Cirac2020, Orus2014} is also known as einsum, factor graph or sum-product network~\cite{Bishop2006} in different contexts,
it can be viewed as a generalization of matrix multiplication to multiple tensors contraction.
People often use Einstein's notation to represent a tensor network, e.g. the matrix multiplication between two matrices $A$ and $B$ can be represented in Einstein's notation as $C_{ik} = A_{ij}B_{jk}$,
where we use a label in subscripts to represent a degree of freedom.
We enumerate these degree of freedoms and accumulate the product of tensor elements to the output tensor.
In the standard notation of Einstein's summation or tensor network in physics, each index appears precisely twice.
Hence people can represent a tensor network as a simple graph,
where a tensor is a vertex and two tensors sharing the same label are connected.
In the main text, we do not have this restriction. An index can appear arbitrary times.
The graphical representation of our generlized tensor network is hypergraph, in which an edge (label) can be shared by an arbitrary number of vertices (tensors).
These two notations are equivalent in representation power because one can easily translate a generalized tensor network to the standard notation by adding $\delta$ tensors - a high dimensional equivalence of identity matrix.
However, introducing $\delta$ tensors can potentially increase the contraction complexity a lot. We illustrate this point in \App{app:tensorbad}.
\blue{Is it better to introduce the contraction orders here or in example~\ref{eg:tensorcontraction}?}

@@

\begin{figure} 
    \centering
    \includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm}, clip]{figures/fig1.pdf}
    \caption{Benchmark results for computing different properties of independent sets of a random three regular graph with different tensor element types.
    The time in these plots only includes tensor network contraction, without taking the contraction order finding and just in time compiling time into account.
    Legends are properties, algebra and devices that we used in the computation, one can find the corresponding computed property in \Tbl{tbl:generictypes}.
    (a) time and space complexity versus the number of vertices for the benchmarked graphs.
    (b) The computing time for calculating the MIS size and for counting of the number of independent sets (ISs), the number of MISs, and the number of independent sets having size $\alpha(G)$ and $\alpha(G)-1$ (denoted as ``MISs and (MIS-1)s'').
    (c) The computing time for calculating the independence polynomials with different approaches.
    (d) The computing time for configuration enumeration, including the enumeration of all independent set configurations, a single MIS configuration, all MIS configurations, all independent set configurations having size $\alpha(G)$ and $\alpha(G)-1$,  with or without bounding the enumeration space.
    \red{suggestions for the figure: 1. the legend seems to overlap the lines, especially the large subfigure. Not sure what's a better way to include the legend there. 2. I see you guys don't typically use markers. We typically use markers to distinguish the different lines, as well as using the color. Especially, for some journals, there is a physical non-color copy, so some requires the figures to be understandable without colors.}
    }
    \label{fig:benchmark}
\end{figure}

Figure (a) shows the time and space complexity (without slicing) of tensor network contraction for variaty of graph sizes, where the space complexity is the same as the tree width of the problem graph.
It is obtained with tensor network contraction order optimization algorithm in Ref.~\cite{Kalachev2021}.
In practice, slicing technique is used to fit the computation into a 32GB memory, the target tree width of slicing is $27$.
One can see all the computing times in figure (b), (c) and (d) have a strong correlation with the treewidth.
%The computations shown in \Fig{fig:benchmark} include independent set counting, finding maximum independent set size (panel b),
%computing independence polynomial (panel c) and find one independent set and enumerating independent sets (panel d).
Among these benchmarks, computational tasks with data types \texttt{T (CPU)},
\texttt{$\mathbbm{R}$ (CPU)}, \texttt{$\mathbbm{R}$ (GPU)}, \texttt{$\mathbbm{C}$ (CPU)}, \texttt{$\mathbbm{C}$ (GPU)} and \texttt{T+bounding (CPU)} can utilize fast BLAS functions, hence 
are much faster comparing to non-BLAS methods in the same category.
GPU computes much faster than CPU in all cases when the problem scale is large enough so that the actual computing time is comparable or larger than the launching time of CUDA kernels.
Most algebras can be computed on GPU, except those requiring dynamic sized structures, i.e. \texttt{PN} and \texttt{SN}.
In figure (c), one can see the Fourier transformation based method is the fastest in computing the independence polynomial,
however, it maybe suffer from the round off errors. The finite field (GF$(p)$) approach is the only method that does not incur round-off errors and can run on a GPU.
In figure (d), one can see the technique to bound the enumeration space improves the performance for more than one order in enumerating the MISs.
Bounding can also reduce the memory usage significantly, without which the largest computable graph size is only $\sim150$ on a 32GB memory device.

@@

\section{Discussion and conclusion}
In this paper, we introduce an abstract algebra formalism to compute properties of independent sets.
The properties include the MIS size, number of independent set of a given size, and enumeration of independent sets of a give size.
For each property, we design a algebra being a commutative semiring, and map the property computation to the contraction of a tensor network with such element type.
%We derived the backward rule for tropical tensor network to bound the search of solution space.
%Although many of these properties are global, we can encode them to different tensor element types as commutative semirings.
We call this method generic programming tensor network, and its power is not limited to the independent set problem.
In \App{app:otherproblems}, we show how to map the maximal independent set problem, matching problem, $k$-coloring problem, max cut problem and set packing problem to tensor networks.
Moreover, since the independence polynomial is closely related to the matching polynomial~\cite{Levit2005}, the clique polynomial~\cite{Hoede1994}, and the vertex cover polynomial~\cite{Akbari2013},
our algorithm to compute the independence polynomial can also be used to compute these graph polynomials.
%Some properties of the independence polynomial, such as the uni-modality, log-concavity, and roots of the polynomial, are well studied in the literature for special classes of graphs~\cite{Levit2005}.
%We hope our algorithms to calculate the independence polynomial can also help further research in these directions.
We show some of the Julia language implementations in Appendix~\ref{sec:technical} and you will find it surprisingly short.
A complete implementation can be found in our Github repository~\cite{GraphTensorNetworks}.
%What we need to do is just defining two operations $\oplus$ and $\odot$ and two special elements $\mymathbb{0}$ and $\mymathbb{1}$.
%The style that we program is called generic programming,

@@

We can modify the tensor network for computing the independence polynomial to include this restriction. Instead of defining the restriction on vertices and edges, it is more natural to define it on $N[v]$:
\begin{equation}\label{eq:maximal}
    T(x_v)_{s_1,s_2,\ldots,s_{|N(v)|},s_v} = \begin{cases}
        s_vx_v & s_1=s_2=\ldots=s_{|N(v)|}=0,\\
        1-s_v& \text{otherwise}.\\
    \end{cases}
\end{equation}
Intuitively, it means if all the neighborhood vertices are not in $I_{m}$, i.e., $ s_1=s_2=\ldots=s_{|N(v)|}=0$, then $v$ should be in $I_{m}$ and contribute a factor $x_{v}$,
otherwise, if any of the neighborhood vertices is in $I_{m}$, then $v$ cannot be in $I_{m}$.
As an example, for a vertex of degree 2, the resulting rank-3 tensor is
\begin{equation}
    T(x_v)=\left(\begin{matrix}
    \left(\begin{matrix}
        0 &1 \\
        1 &1
    \end{matrix}\right)\\
    \left(\begin{matrix}
        x_v &0 \\
        0 &0
    \end{matrix}\right)
    \end{matrix}\right).
\end{equation}

@@

One can see that the average degree of a tensor is increased.
The computational complexity of this new tensor network contraction is often greater than the one for computing the independence polynomial.
However, for most sparse graphs, this tensor network contraction approach is still much faster than enumerating all the maximal cliques on its complement graph using the Bron-Kerbosch algorithm~\cite{Bron1973}, which is the standard algorithm that we are aware of to compute the maximal independence polynomial.
We show the benchmark of computing the maximal independent set properties in \Fig{fig:benchmark-maximal},
including a comparison to the Bron-Kerbosch algorithm from Julia package Graphs~\cite{Graphs}.
%Since the network for computing maximal independent set properties is different from the one for computing the independence polynomial,
the tree width of this tensor network is significantly larger, hence only small graphs can be benchmarked.
The time for the tensor netwrok approach and the Bron-Kerbosch approach to enumerate all maximal independent sets are comparable,
while the tensor network does counting much more efficiently.
Due to the memory limit, this Bron-Kerbosch algorithm stops working at size $70$ and above.

@@

\subsection{k-Coloring}
Let us use 3-coloring problem defined on vertices as an example.
For a vertex $v$, we define the degree of freedoms $c_v\in\{1,2,3\}$ and a vertex tensor labeled by it as
\begin{equation}
    W(v) = \left(\begin{matrix}
        r_v\\
        g_v\\
        b_v
    \end{matrix}\right).
\end{equation}
For an edge $(u, v)$, we define an edge tensor as a matrix labeled by $(c_u, c_v)$ to specify the constraint
\begin{equation}
    B = \left(\begin{matrix}
        0 & 1 & 1\\
        1 & 0 & 1\\
        1 & 1 & 0
    \end{matrix}\right).
\end{equation}
The number of possible coloring can be obtained by contracting this tensor network by setting vertex tensor elements $r_v, g_v$ and $b_v$ to $1$.
By designing generic types as tensor elements, one can get other properties.
Similarly, one can define the k-coloring problem on edges too by switching the roles of edges and vertices.

\subsection{Max cut problem}
Max cut problem is also known as the boolean spinglass problem.
For a vertex $v\in V$, we define a boolean degree of freedom $s_v\in\{0, 1\}$.
Then the max cut problem can be encoded to tensor networks by mapping an edge $(i,j)\in E$ to an edge matrix labeled by $s_is_j$
\begin{equation}
    B(x_{\langle i, j\rangle}) = \left(\begin{matrix}
        1 & x_{\langle i, j\rangle}\\
        x_{\langle i, j\rangle} & 1
    \end{matrix}\right),
\end{equation}
where variable $x_{\langle i, j\rangle}$ represents a cut on edge $(i, j)$ or a domain wall of an Ising spinglass.
Similar to other problems, we can define a polynomial about edges variables by setting $x_{\langle i, j\rangle} = x$,
where its $k$th coefficient is two times the number of configurations of cut size $k$.

>>>>

\section{Introduction}
In graph theory and combinatorial optimization, there are many interesting and hard computational problems concerning various properties of independent sets.
%An independent set of a graph is a subset of its vertices that no two vertices in this set are adjacent to each other.
For an undirected graph $G = (V,E)$, an independent set $I \subseteq V$ is a set of vertices that for any vertex pair $u,v \in I$, $(u, v) \not\in E$. %there is no edge connecting $u$ and $v$ in $G$. 
One of the independent set properties that many computational complexity theorists are most interested in is to find a maximum independent set (MIS) and the size of the MIS, $\alpha(G) \equiv \max_{I}|I|$. 
\blue{The primary interest of many computational complexity theorists is finding the maximum size of such vertex sets $\alpha(G) \equiv \max_{I}|I|$ and one of the sets of this size.}
Finding $\alpha(G)$ exactly is a hard computational problem, and it is NP-hard even to approximate $\alpha(G)$ within $|V|^{1-\epsilon}$ ~\cite{Hastad1996},
meaning there is unlikely a polynomial time algorithm to find an independent set with size $\alpha(G)/|V|^{1-\epsilon}$ for an arbitrarily small positive $\epsilon$.
Naive exhaustive search for an MIS requires a computational time of $O\left(2^{|V|} \right)$.
More efficient exact algorithms have been developed for finding MISs, such as the branching algorithms~\cite{Tarjan1977, Robson1986} and dynamic programming based algorithms~\cite{Courcelle1990, Fomin2013}.
The branching algorithms can reduce the base of the exponential time scaling to, e.g., $1.1996^{|V|} {|V|}^{O(1)}$~\cite{Xiao2017}, 
while dynamic programming approaches~\cite{Courcelle1990, Fomin2013} works better for graphs with a small treewidth $\text{tw}(G)$, producing algorithms of complexity $O(2^{\text{tw}(G)} \text{tw}(G) |V|)$.
There is immense interest in finding better algorithms to solve the MIS problem, not only because it has 
a wide range of applications in scheduling, logistics, wireless networks and telecommunication, and computer vision, etc.~\cite{Butenko2003, Wu2015}, 
but also because it is a well-known NP-complete problem that can be mapped from many other important combinatorial optimization problems in polynomial time such as the 3-satisfiability problem, the maximum clique problem, and the minimum vertex cover problem~\cite{Moore2011}.

%more concretely, an maximal independent set of a graph $G$ corresponds to a maximum clique in the complement graph of $G$,
%and the complement vertex set of an MIS corresponds to a minimum vertex cover of the graph $G$.
%In this paper, the term ``maximal'' has a different meaning from ``maximum''  which we discussed above; a maximal independent set is an independent set that is not a subset of any other independent set, but its size may not be the maximum.
\blue{We need to distinguish ourself in the problem to solve.}
In this paper, we do not limit our study to just finding an MIS and its size. 
We introduce a tensor-network based framework to compute the various \textit{properties} pertaining to independent sets.
These properties include, for example, the number and enumeration of independent sets or maximal independent sets of a given size.
We map these independent set problems into generic tensor network contraction with specially designed tensor element algebra.
% There are many other interesting problems pertaining to independent sets, such as the number of independent sets of a given size
% and enumerating independent sets at a give size or what we call computing \textit{properties}.
Some of these problems are of great interest in physics applications such as the hard-core lattice gas model~\cite{Dyre2016, Fernandes2007}
in statistical mechanics and the Rydberg hamiltonian with neutral atoms~\cite{Pichler2018, Ebadi2022} \red{replace with experiment paper when ready};
they can, for example, be used to understand phase transitions, to identify harder graphs in an ensemble of graphs~\cite{Ebadi2022} and to analyze the overlap gap property~\cite{Gamarnik2013, Gamarnik2019}.

@@

\section{Tensor networks}
Tensor network~\cite{Cirac2020, Orus2014} is also known as einsum, factor graph or sum-product network~\cite{Bishop2006} in different contexts,
it can be viewed as a generalization of matrix multiplication to multiple tensors contraction.
People often use Einstein's notation to represent a tensor network, e.g. the matrix multiplication between two matrices $A$ and $B$ can be represented in Einstein's notation as $C_{ik} = A_{ij}B_{jk}$,
where we use a label in subscripts to represent a degree of freedom.
We enumerate these degree of freedoms and accumulate the product of tensor elements to the output tensor.
In the standard notation of Einstein's summation or tensor network in physics, each index appears precisely twice.
Hence people can represent a tensor network as a simple graph,
where a tensor is a vertex and two tensors sharing the same label are connected.
In the main text, we do not have this restriction. An index can appear arbitrary times.
The graphical representation of our generalized tensor network is hypergraph, in which an edge (label) can be shared by an arbitrary number of vertices (tensors).
These two notations are equivalent in representation power because one can easily translate a generalized tensor network to the standard notation by adding $\delta$ tensors - a high dimensional equivalence of identity matrix.
However, introducing $\delta$ tensors can potentially increase the contraction complexity a lot. We illustrate this point in \App{app:tensorbad}.
\blue{Is it better to introduce the contraction orders here or in example~\ref{eg:tensorcontraction}?}

@@

\begin{figure} 
    \centering
    \includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm}, clip]{figures/fig1.pdf}
    \caption{Benchmark results for computing different properties of independent sets of a random three regular graph with different tensor element types.
    The time in these plots only includes tensor network contraction, without taking the contraction order finding and just in time compiling time into account.
    Legends are properties, algebra and devices that we used in the computation, one can find the corresponding computed property in \Tbl{tbl:generictypes}.
    (a) time and space complexity versus the number of vertices for the benchmarked graphs.
    (b) The computing time for calculating the MIS size and for counting of the number of independent sets (ISs), the number of MISs, and the number of independent sets having size $\alpha(G)$ and $\alpha(G)-1$ (denoted as ``MISs and (MIS-1)s'').
    (c) The computing time for calculating the independence polynomials with different approaches.
    (d) The computing time for configuration enumeration, including the enumeration of all independent set configurations, a single MIS configuration, all MIS configurations, all independent set configurations having size $\alpha(G)$ and $\alpha(G)-1$,  with or without bounding the enumeration space.
    \red{suggestions for the figure: 1. the legend seems to overlap the lines, especially the large sub-figure. Not sure what's a better way to include the legend there. 2. I see you guys don't typically use markers. We typically use markers to distinguish the different lines, as well as using the color. Especially, for some journals, there is a physical non-color copy, so some requires the figures to be understandable without colors.}
    }
    \label{fig:benchmark}
\end{figure}

Figure (a) shows the time and space complexity (without slicing) of tensor network contraction for variety of graph sizes, where the space complexity is the same as the tree width of the problem graph.
It is obtained with tensor network contraction order optimization algorithm in Ref.~\cite{Kalachev2021}.
In practice, slicing technique is used to fit the computation into a 32GB memory, the target tree width of slicing is $27$.
One can see all the computing times in figure (b), (c) and (d) have a strong correlation with the tree width.
%The computations shown in \Fig{fig:benchmark} include independent set counting, finding maximum independent set size (panel b),
%computing independence polynomial (panel c) and find one independent set and enumerating independent sets (panel d).
Among these benchmarks, computational tasks with data types \texttt{T (CPU)},
\texttt{$\mathbbm{R}$ (CPU)}, \texttt{$\mathbbm{R}$ (GPU)}, \texttt{$\mathbbm{C}$ (CPU)}, \texttt{$\mathbbm{C}$ (GPU)} and \texttt{T+bounding (CPU)} can utilize fast BLAS functions, hence 
are much faster comparing to non-BLAS methods in the same category.
GPU computes much faster than CPU in all cases when the problem scale is large enough so that the actual computing time is comparable or larger than the launching time of CUDA kernels.
Most algebras can be computed on GPU, except those requiring dynamic sized structures, i.e. \texttt{PN} and \texttt{SN}.
In figure (c), one can see the Fourier transformation based method is the fastest in computing the independence polynomial,
however, it maybe suffer from the round off errors. The finite field (GF$(p)$) approach is the only method that does not incur round-off errors and can run on a GPU.
In figure (d), one can see the technique to bound the enumeration space improves the performance for more than one order in enumerating the MISs.
Bounding can also reduce the memory usage significantly, without which the largest computable graph size is only $\sim150$ on a 32GB memory device.

@@

\section{Discussion and conclusion}
In this paper, we introduce an abstract algebra formalism to compute properties of independent sets.
The properties include the MIS size, number of independent set of a given size, and enumeration of independent sets of a give size.
For each property, we design a algebra being a commutative semiring, and map the property computation to the contraction of a tensor network with such element type.
%We derived the backward rule for tropical tensor network to bound the search of solution space.
%Although many of these properties are global, we can encode them to different tensor element types as commutative semirings.
We call this method generic programming tensor network, and its power is not limited to the independent set problem.
In \App{app:otherproblems}, we show how to map the maximal independent set problem, matching problem, $k$-colouring problem, max cut problem and set packing problem to tensor networks.
Moreover, since the independence polynomial is closely related to the matching polynomial~\cite{Levit2005}, the clique polynomial~\cite{Hoede1994}, and the vertex cover polynomial~\cite{Akbari2013},
our algorithm to compute the independence polynomial can also be used to compute these graph polynomials.
%Some properties of the independence polynomial, such as the uni-modality, log-concavity, and roots of the polynomial, are well studied in the literature for special classes of graphs~\cite{Levit2005}.
%We hope our algorithms to calculate the independence polynomial can also help further research in these directions.
We show some of the Julia language implementations in Appendix~\ref{sec:technical} and you will find it surprisingly short.
A complete implementation can be found in our Github repository~\cite{GraphTensorNetworks}.
%What we need to do is just defining two operations $\oplus$ and $\odot$ and two special elements $\mymathbb{0}$ and $\mymathbb{1}$.
%The style that we program is called generic programming,

@@

We can modify the tensor network for computing the independence polynomial to include this restriction. Instead of defining the restriction on vertices and edges, it is more natural to define it on $N[v]$:
\begin{equation}\label{eq:maximal}
    T(x_v)_{s_1,s_2,\ldots,s_{|N(v)|},s_v} = \begin{cases}
        s_vx_v & s_1=s_2=\ldots=s_{|N(v)|}=0,\\
        1-s_v& \text{otherwise}.\\
    \end{cases}
\end{equation}
Intuitively, it means if all the neighbourhood vertices are not in $I_{m}$, i.e., $ s_1=s_2=\ldots=s_{|N(v)|}=0$, then $v$ should be in $I_{m}$ and contribute a factor $x_{v}$,
otherwise, if any of the neighbourhood vertices is in $I_{m}$, then $v$ cannot be in $I_{m}$.
As an example, for a vertex of degree 2, the resulting rank-3 tensor is
\begin{equation}
    T(x_v)=\left(\begin{matrix}
    \left(\begin{matrix}
        0 &1 \\
        1 &1
    \end{matrix}\right)\\
    \left(\begin{matrix}
        x_v &0 \\
        0 &0
    \end{matrix}\right)
    \end{matrix}\right).
\end{equation}

@@

One can see that the average degree of a tensor is increased.
The computational complexity of this new tensor network contraction is often greater than the one for computing the independence polynomial.
However, for most sparse graphs, this tensor network contraction approach is still much faster than enumerating all the maximal cliques on its complement graph using the Bron-Kerbosch algorithm~\cite{Bron1973}, which is the standard algorithm that we are aware of to compute the maximal independence polynomial.
We show the benchmark of computing the maximal independent set properties in \Fig{fig:benchmark-maximal},
including a comparison to the Bron-Kerbosch algorithm from Julia package Graphs~\cite{Graphs}.
%Since the network for computing maximal independent set properties is different from the one for computing the independence polynomial,
the tree width of this tensor network is significantly larger, hence only small graphs can be benchmarked.
The time for the tensor network approach and the Bron-Kerbosch approach to enumerate all maximal independent sets are comparable,
while the tensor network does counting much more efficiently.
Due to the memory limit, this Bron-Kerbosch algorithm stops working at size $70$ and above.

@@

\subsection{k-Colouring}
Let us use 3-colouring problem defined on vertices as an example.
For a vertex $v$, we define the degree of freedoms $c_v\in\{1,2,3\}$ and a vertex tensor labelled by it as
\begin{equation}
    W(v) = \left(\begin{matrix}
        r_v\\
        g_v\\
        b_v
    \end{matrix}\right).
\end{equation}
For an edge $(u, v)$, we define an edge tensor as a matrix labelled by $(c_u, c_v)$ to specify the constraint
\begin{equation}
    B = \left(\begin{matrix}
        0 & 1 & 1\\
        1 & 0 & 1\\
        1 & 1 & 0
    \end{matrix}\right).
\end{equation}
The number of possible colouring can be obtained by contracting this tensor network by setting vertex tensor elements $r_v, g_v$ and $b_v$ to $1$.
By designing generic types as tensor elements, one can get other properties.
Similarly, one can define the k-colouring problem on edges too by switching the roles of edges and vertices.

\subsection{Max cut problem}
Max cut problem is also known as the boolean spin glass problem.
For a vertex $v\in V$, we define a boolean degree of freedom $s_v\in\{0, 1\}$.
Then the max cut problem can be encoded to tensor networks by mapping an edge $(i,j)\in E$ to an edge matrix labelled by $s_is_j$
\begin{equation}
    B(x_{\langle i, j\rangle}) = \left(\begin{matrix}
        1 & x_{\langle i, j\rangle}\\
        x_{\langle i, j\rangle} & 1
    \end{matrix}\right),
\end{equation}
where variable $x_{\langle i, j\rangle}$ represents a cut on edge $(i, j)$ or a domain wall of an Ising spin glass.
Similar to other problems, we can define a polynomial about edges variables by setting $x_{\langle i, j\rangle} = x$,
where its $k$th coefficient is two times the number of configurations of cut size $k$.
